03 Data Vizualization

### 2022-12-17 03_data_viz.Rmd
supp figures

```{r mean_generation_time}
p_gen_time_mean <- turnover_weighted %>% 
  select(c(terrain, inc_time_d, gen_d_weighted, gen_d_mc_weighted)) %>% 
  pivot_longer(cols = c(gen_d_weighted, gen_d_mc_weighted),
               values_to = "gen_d",
               names_to = "type") %>% 
  filter(inc_time_d == 7) %>% 
  mutate(terrain = case_when(
    terrain == "Gordon Gulch Conifer Forest" ~ "Conifer Forest",
    terrain == "Marshall Mesa Grassland" ~ "Grassland",
    terrain == "Niwot Ridge Tundra" ~ "Tunrda" )) %>% 
  # group_by(terrain, inc_time_d) %>% 
  # summarize(terrain = terrain,
  #           gen_d_weighted = gen_d_weighted) %>% 
  # unique() %>% 
  # Plot it
  ggplot(aes(x = terrain,
             y = gen_d,
             fill = type)) +
  geom_col(width = 0.5,
           position = "dodge") +
  scale_fill_tron() +
  labs(x = "Soil Type",
       y = "Weighted Mean Generation Time (d)") +
  theme_classic() +
  theme()
p_gen_time_mean

cowplot::save_plot(
  plot = p_gen_time_mean,
  filename = "fig_output/generation_time_mean_mc.pdf",
  base_height = 4,
  base_width = 7
)
```

@@

  ggplot(
    aes(
      x = inc_time_d,
      y = dF * 10000, # convert from at% to ppm
      group = compound,
      color = terrain,
      fill = terrain,
      alpha = type,
      size = type
    )
  ) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(terrain)) +
  scale_color_manual(values = terrain_color_swatch) +
  scale_fill_manual(values = terrain_fill_swatch) +
  scale_alpha_discrete(range = c(0.2, 1)) +
  scale_size_discrete(range = c(1.5, 2)) +
  scale_x_continuous(breaks = c(0, 1,2,3,4,5,6,7)) +
  theme_classic() +
  labs(
    x = "Time elapsed (days)",
    y = TeX("$ \\Delta ^2H$ (ppm)")
  ) +
  theme(legend.position = "none",
        strip.background = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 10),
        panel.grid.major.y = element_line(linetype = "dotted"))
p_dF

cowplot::save_plot(
  filename = "fig_output/p_dF.pdf",
  plot = p_dF,
  base_width = 11,
  base_height = 4
  )

@@

turnover_fit_tidy <- broom::tidy(lm(data = turnover_cleaned, u_d_mc  ~ compound_ug_per_g_soil))
turnover_fit_glance <- broom::glance(lm(data = turnover_cleaned, u_d_mc  ~ compound_ug_per_g_soil))
writexl::write_xlsx(turnover_fit_tidy, path = "data_output/turnover_fit_lm.xlsx")
writexl::write_xlsx(turnover_fit_glance, path = "data_output/turnover_fit_lm_stats.xlsx")

p_abundance_v_turnover2 <- turnover %>% 
  mutate(
    terrain = case_when(
      terrain == "Gordon Gulch Conifer Forest" ~ "Conifer Forest",
      terrain == "Marshall Mesa Grassland" ~ "Grassland",
      terrain == "Niwot Ridge Tundra" ~ "Tundra"
  )) %>% 
  filter(inc_time_d != 0) %>% 
  ggplot(aes(
    x = compound_ug_per_g_soil,
    y = u_d_mc
  )) +
  # Individual data points
  geom_point(
    aes(color = terrain),
    alpha = 0.6,
    ) +
  geom_polygon(
    data = hull_data,
    aes(
       fill = terrain,
       color = terrain
       ),
     show.legend = FALSE,
     alpha = 0.15,
     size = 0.15
    ) +
  geom_smooth(
    method = "lm",
    formula = "y~x",
    se = FALSE,
    color = "gray",
    size = 0.3,
    linetype = "dotted"
  ) +
  scale_color_manual(values = terrain_color_swatch) +
  theme_bw() +
  labs(
    x = "µg analyte (per g soil)",
    y = TeX("Compound-specific growth rate (day$^{-1}$)"),
    size = TeX("$ \\Delta ^2H$ (ppm)"),
    color = " "
  ) +
  theme(
    strip.background = element_blank(),
    legend.position = c(0.9, 0.55),
    legend.background = element_blank(),
    panel.grid = element_blank(),
  )
p_abundance_v_turnover2

cowplot::save_plot(filename = "fig_output/abundance_v_turnover_2p.pdf",
                   plot = p_abundance_v_turnover2,
                   base_height = 6,
                   base_width = 8)

@@

```{r}
p_abundance_v_turnover_w_inset <- ggdraw(p_abundance_v_turnover2) +
  draw_plot(
    p_abundance_v_turnover_inset, 
    x = 0.63, 
    y = 0.63, 
    width = 0.35, 
    height = 0.35
    )
cowplot::save_plot(filename = "fig_output/abundance_v_turnover_w_inset.pdf",
                   plot = p_abundance_v_turnover_w_inset,
                   base_height = 6,
                   base_width = 8)
```

@@

>>>>

```{r mean_generation_time}
p_gen_time_mean <- turnover_weighted %>% 
  select(c(terrain, inc_time_d, gen_d_weighted, gen_d_mc_weighted)) %>% 
  pivot_longer(cols = c(gen_d_weighted, gen_d_mc_weighted),
               values_to = "gen_d",
               names_to = "type") %>% 
  filter(inc_time_d == 7) %>% 
  mutate(terrain = case_when(
    terrain == "Gordon Gulch Conifer Forest" ~ "Conifer Forest",
    terrain == "Marshall Mesa Grassland" ~ "Grassland",
    terrain == "Niwot Ridge Tundra" ~ "Tunrda" )) %>% 
  # group_by(terrain, inc_time_d) %>% 
  # summarize(terrain = terrain,
  #           gen_d_weighted = gen_d_weighted) %>% 
  # unique() %>% 
  # Plot it
  ggplot(aes(x = terrain,
             y = gen_d,
             fill = type)) +
  geom_col(width = 0.5,
           position = "dodge") +
  scale_fill_tron() +
  labs(x = "Soil Type",
       y = "Weighted Mean Generation Time (d)") +
  theme_classic() +
  theme()
p_gen_time_mean
# 
# cowplot::save_plot(
#   plot = p_gen_time_mean,
#   filename = "fig_output/generation_time_mean_mc.pdf",
#   base_height = 4,
#   base_width = 7
# )
```
## Memory effect on u_d

@@

  ggplot(
    aes(
      x = inc_time_d,
      y = dF * 10000, # convert from at% to ppm
      group = compound,
      color = terrain,
      fill = terrain,
      alpha = type,
      size = type
    )
  ) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(terrain)) +
  scale_color_manual(values = terrain_color_swatch) +
  scale_fill_manual(values = terrain_fill_swatch) +
  scale_alpha_discrete(range = c(0.2, 1)) +
  scale_size_discrete(range = c(1.5, 2)) +
  scale_x_continuous(breaks = c(0, 1,2,3,4,5,6,7)) +
  theme_classic() +
  labs(
    x = "Time elapsed (days)",
    y = TeX("$ \\Delta ^2H$ (ppm)")
  ) +
  theme(legend.position = "none",
        strip.background = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 10),
        panel.grid.major.y = element_line(linetype = "dotted"))
p_dF
# 
# cowplot::save_plot(
#   filename = "fig_output/p_dF.pdf",
#   plot = p_dF,
#   base_width = 11,
#   base_height = 4
#   )
#   
```

@@

# Calculate summary statistics and export them
turnover_fit_tidy <- broom::tidy(lm(data = turnover_cleaned, u_d_mc  ~ compound_ug_per_g_soil))
turnover_fit_glance <- broom::glance(lm(data = turnover_cleaned, u_d_mc  ~ compound_ug_per_g_soil))

turnover_corr_test <- broom::tidy(
  cor.test(
    x = turnover_cleaned$compound_ug_per_g_soil,
    y = turnover_cleaned$u_d_mc,
    method = "pearson"
  )
  )

writexl::write_xlsx(turnover_fit_tidy, path = "data_output/turnover_fit_lm.xlsx")
writexl::write_xlsx(turnover_fit_glance, path = "data_output/turnover_fit_lm_stats.xlsx")
writexl::write_xlsx(turnover_corr_test, path = "data_output/turnover_corr_test.xlsx")

# Plot it with the linear model!
p_abundance_v_turnover2 <- turnover %>% 
  mutate(
    terrain = case_when(
      terrain == "Gordon Gulch Conifer Forest" ~ "Conifer Forest",
      terrain == "Marshall Mesa Grassland" ~ "Grassland",
      terrain == "Niwot Ridge Tundra" ~ "Tundra"
  )) %>% 
  filter(inc_time_d != 0) %>% 
  ggplot(aes(
    x = compound_ug_per_g_soil,
    y = u_d_mc
  )) +
  # Individual data points
  geom_point(
    aes(color = terrain),
    alpha = 0.6,
    ) +
  geom_polygon(
    data = hull_data,
    aes(
       fill = terrain,
       color = terrain
       ),
     show.legend = FALSE,
     alpha = 0.15,
     size = 0.15
    ) +
  geom_smooth(
    method = "lm",
    formula = "y~x",
    se = FALSE,
    color = "gray",
    size = 0.3,
    linetype = "dotted"
  ) +
  scale_color_manual(values = terrain_color_swatch) +
  theme_bw() +
  labs(
    x = "µg analyte (per g soil)",
    y = TeX("Compound-specific growth rate (day$^{-1}$)"),
    size = TeX("$ \\Delta ^2H$ (ppm)"),
    color = " "
  ) +
  theme(
    strip.background = element_blank(),
    legend.position = c(0.9, 0.55),
    legend.background = element_blank(),
    panel.grid = element_blank(),
  )
p_abundance_v_turnover2

# 
# cowplot::save_plot(filename = "fig_output/abundance_v_turnover_2p.pdf",
#                    plot = p_abundance_v_turnover2,
#                    base_height = 6,
#                    base_width = 8)

@@

```{r}
p_abundance_v_turnover_w_inset <- ggdraw(p_abundance_v_turnover2) +
  draw_plot(
    p_abundance_v_turnover_inset, 
    x = 0.63, 
    y = 0.63, 
    width = 0.35, 
    height = 0.35
    )
p_abundance_v_turnover_w_inset

# cowplot::save_plot(filename = "fig_output/abundance_v_turnover_w_inset.pdf",
#                    plot = p_abundance_v_turnover_w_inset,
#                    base_height = 6,
#                    base_width = 8)
```

@@

## SF: µ vs. 1/B

### Compound-specific

```{r}
turnover_cleaned %>% 
  ggplot(
    aes(
    x = 1/compound_ug_per_g_soil,
    y = u_d_mc,
    color = terrain
    )
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(
    method = "lm",
    formula = "y~x",
    se = FALSE
  ) +
  scale_color_manual(values = terrain_color_swatch) +
  labs(
    x = "1/µg analyte per g soil",
    y = "µ (days)"
  ) +
  theme_bw()
```

### Site-specific

```{r}
total_biomass <- turnover_cleaned %>% 
  group_by(terrain) %>% 
  summarize(
    total_ug_per_g_soil = sum(compound_ug_per_g_soil, na.rm = TRUE)
  )

turnover_weighted_w_totals <- turnover_weighted %>% 
  filter(inc_time_d == 7) %>% 
  mutate(
    terrain = case_when(
      terrain == "Gordon Gulch Conifer Forest" ~ "Conifer Forest",
      terrain == "Marshall Mesa Grassland" ~ "Grassland",
      terrain == "Niwot Ridge Tundra" ~ "Tundra"
  )) %>% 
  left_join(total_biomass, by = "terrain")

p_turnover_totals <- turnover_weighted_w_totals %>% 
  ggplot(
    aes(
      x = total_ug_per_g_soil,
      y = u_d_mc_weighted
    )
  ) +
  geom_point(aes(color = terrain),
             size = 3) +
  geom_smooth(
    method = "lm",
    formula = "y~x",
    se = FALSE,
    linetype = "dotted",
    color = "gray",
    size = 0.5
  ) +
  scale_color_manual(values = terrain_color_swatch) +
  labs(
    x = "B (µg)",
    y = TeX("µ $d^{-1}$"),
    caption = "Where B is total biomass measured at a given site, µg per g soil",
    color = ""
  ) +
  theme_bw() +
  theme(
    legend.position = "NA"
  )
p_turnover_totals
```

```{r}
saveRDS(turnover_weighted_w_totals, file = "cache/turnover_weighted_w_totals.RDS")

F2_draft3 <- cowplot::plot_grid(
  p_abundance_v_turnover_w_inset,
  p_turnover_totals,
  rel_widths = c(2, 1),
  rel_heights = c(2, 1),
  labels = c("A", "B")
)
F2_draft3
```

## SF: Sensitivity Testing

### at2H se

```{r}
LH_SIP_data %>% 
  ggplot(
    aes(
      x = calibrated_at2H_without_memory_se,
      y = calibrated_at2H_with_memory_correction_se,
      color = inc_time_d_str
    )
  ) + 
  geom_point(
    alpha = 0.5
  ) +
  geom_abline(
    slope = 1,
    color = "red")
```

### IRMS instrument error

```{r}
instrument_error <- mean(LH_SIP_data$F_0_at2H_se)
instrument_error_u_d <- calculate_turnover(
  a = 0.7,
  FL = 0.5,
  F0 = 0.00000000000000000000001,
  FT = instrument_error,
  t = 7
)

instrument_error_test <- crossing(
  t = seq(1, 30, by = 0.1),
  FL = c(0.1, 0.25, 0.5, 1, 2.5, 5)
) %>% mutate(
  a = 0.7,
  F0 = 0.00000000000000000000001,
  FT = instrument_error,
  u_d = calculate_turnover(
    a = a,
    FT = FT,
    F0 = F0,
    FL = FL,
    t = t
  )
)

instrument_error_test %>% 
  ggplot(
    aes(
      x = t,
      y = u_d,
      color = as.factor(FL)
    )
  ) +
  geom_line() +
  geom_point(
    y = instrument_error_u_d,
    x = 7,
    size = 2,
    color = "black",
    shape = 1
    ) +
  scale_color_viridis_d() +
  theme_classic() +
  labs(
    x = "Incubation Time (days)",
    y = TeX("Growth Rate Uncertainty (day $^{-1})"),
    color = "Tracer Strength (%)"
  )
```

### IRMS instrument "turnover"

instrument_turnover <- turnover_summarized %>% 
  mutate(
    instrument_u_d = calculate_turnover(
      a = a, # Assimilation efficiency defined in data reduction step
      FT = F_0_at2H_se, # 2F instrument error (at%)
      F0 = 0.0000000000000000000001, # 2F at t0 (at%)
      FL = f_label, # 2F of label (at%)
      t = d_t # change in time
      ),
    instrument_over_actual_u_d = instrument_u_d / u_d_mc
  ) %>%
  filter(inc_time_d == 7) %>% 
  group_by(compound, terrain) %>% 
  summarize(
    instrument_u_d = instrument_u_d,
    u_d_mc = u_d_mc,
    instrument_over_actual_u_d_mn = mean(
      instrument_over_actual_u_d, na.rm = TRUE
      )
    ) %>% 
  filter(
    !is.na(instrument_over_actual_u_d_mn),
    is.finite(instrument_over_actual_u_d_mn)
    ) %>% 
  filter(!compound %in% c("20:0", "22:0", "24:0", "18:0", "cyclo 17:0"))

instrument_frac_error <- mean(instrument_turnover$instrument_over_actual_u_d_mn)

paste0("The accuracy of IRMS is such that we can accurately quantify deuterium incorporation corresponding to ", round(instrument_frac_error *100, 2), "% of an organism's growth rate.")

## Turnover MC effect

```{r}
turnover_summarized %>% 
  ggplot(aes(
    x = u_d,
    y = u_d_mc,
    color = as.factor(inc_time_d)
  )) +
  geom_point() +
  geom_abline(linetype = "dashed") +
  labs(
    x = "Growth Rate (day^-1)",
    y = "Memory Corrected Growth Rate (day^-1)"
  )
```

```{r}
turnover_summarized %>% 
  filter(inc_time_d == 7) %>% 
  pivot_longer(
    cols = c(u_d, u_d_mc),
    names_to = "corrected",
    values_to = "u_d") %>% 
  group_by(compound) %>%
  mutate(y = paste(round(mean(rt), 0), compound)) %>% 
  ungroup() %>% 
  arrange(desc(y)) %>% 
  ggplot(
    aes(
      x = u_d,
      y = y,
      color = corrected,
      size = area2
      )
  ) +
    geom_point() +
  facet_wrap(vars(terrain))
```
